---
title: "Assessing the Impact of Coffee Characteristics on Quality Classification Using GLM"
author: "Group 11"
number-sections: true
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: 
    keep-tex: true
header-includes: |
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
editor_options: 
  chunk_output_type: console
execute:
  echo: true
  echo-default: null
  eval: true
  warning: false
  message: false
---
```{r}
#| echo: false
#| label: libraries
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(tidyr)
library(ggplot2)
library(broom)
library(knitr)
library(dplyr)
library(grid)
library(gridExtra)
library(patchwork)
```
# Introduction {#sec-Intro}

Coffee quality assessment is a crucial aspect of the coffee industry, impacting both market value and consumer preferences. This study aims to explore the factors that influence whether a batch of coffee is classified as "Good" or "Poor", using data from the Coffee Quality Database (CQD).

The dataset includes moren than 1000 coffee samples from different countries and records sensory attributes (aroma, flavor, and acidity), production characteristics (harvest year, altitude), and defect counts. The coffee batches are assigned a quality classification based on an overall score threshold (≥82.5 as "Good", \<82.5 as "Poor").

The primary goal of this study is to analyze the relationship between coffee quality and factors mentioned above, especially the sensory attributes(aroma, flavor, and acidity).

# Exploratory Data Analysis {#sec-EDA}

## Data preparing & Cleaning

```{r}
#data cleaning
data<-read.csv("dataset11.csv")
data<-na.omit(data)
data$Qualityclass_dummy<-ifelse(data$Qualityclass=="Good",1,0) #for "Good"=1 "Poor"=0
data$Qualityclass <- as.factor(data$Qualityclass)
data$harvested <- as.factor(data$harvested)
#Outlier in altitude_mean_meters
sum(data$altitude_mean_meters>8848)+sum(data$altitude_mean_meters<0)
data=data%>%
  filter(data$altitude_mean_meters<8848 & data$altitude_mean_meters>0)
#standarized for altitude_mean_meters
data$altitude_mean_meters=data$altitude_mean_meters/1000
```

```{r}
#| label: fig-1
#| fig-cap: the density of explanatory variables
#| fig-width: 6
#| fig-height: 4.5
#| fig-align: center
par(mfrow=c(2,2))
plot(density(data$aroma), col = "blue", main = "Density Plot of aroma")
plot(density(data$flavor), col = "blue", main = "Density Plot of flavor")
plot(density(data$acidity), col = "blue", main = "Density Plot of acidity")
plot(density(data$altitude_mean_meters), col = "blue", main = "Density Plot of altitude_mean_meters")
```

In this dataset, aroma,flavor,acidity and altitude_mean_meters and category_two_defects are continuous variables, Qualityclass and harvested are categorical variables. Through the four density plot, it shows that the first three plot nearly obey Normal distribution, but the plot of altitudes ..( complement ). Then we decide to scale the variable of altitudes.

## Graphical Summaries

```{r}
#change formula
data_long <- data %>%
  pivot_longer(cols = c(aroma,flavor,acidity,category_two_defects,
                        altitude_mean_meters),
               names_to = "Variable",
               values_to = "Value")
```

```{r}
#| label: fig-2
#| fig-cap: the boxplot of continuous explanatory variables 
ggplot(data = data_long, aes(x = Qualityclass, y = Value, fill = Qualityclass))+
  geom_boxplot() +
  facet_wrap(~Variable, scales = "free_y") +  
  theme_minimal() +
  labs(x = "qualityclass",
       y = "Value") +
  theme(legend.position = "none")
```

```{r}
#| label: fig-3
#| fig-cap: the boxplot of continuous explanatory variables
ggplot(data, aes(x=harvested ,  y = ..prop.., group=Qualityclass,
                 fill=Qualityclass)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion")
```

-   The four boxplot

-   The plot of harvested...

## Numerical Summaries

```{r}
#summary
#| label: tbl-1
#| tbl-cap: the summary of dataset 
summary(data)
#####need to add a table for summary use gt()
```

# Formal Data Analysis {#sec-FDA}

$$\ln\left(\frac{p}{1-p}\right) = \alpha + \beta_1 \cdot \textrm{aroma} +
    \beta_2 \cdot \textrm{flavor}+
    \beta_3 \cdot \textrm{acidity}+ 
    \beta_4 \cdot \textrm{defects}+
    \beta_5 \cdot \textrm{altitudes}+
    \beta_6 \cdot \textrm{harvested}$$

-   (each variable need to be explain)

```{r}
#model fitted for original
model1 <- glm(Qualityclass_dummy ~ aroma+flavor+acidity+category_two_defects+altitude_mean_meters+harvested, data = data, family = binomial(link = "logit"))
summary(model1)
summ(model1)
```

-   It shows p-value of category_two_defects and harvested are higher than 0.05

```{r}
#del category_two_defects variable
model2 <- glm(Qualityclass_dummy ~ aroma+flavor+acidity+altitude_mean_meters+harvested, data = data, family = binomial(link = "logit"))
summary(model2)
summ(model2)
```

-   AIC decreased

```{r}
#del harvested variable
model3 <- glm(Qualityclass_dummy ~ aroma+flavor+acidity+altitude_mean_meters, data = data, family = binomial(link = "logit"))
summary(model3)
summ(model3)
```

-   AIC decreased

-   We see that the coefficient for....

```{r}
#del altitude_mean_meters variable
model4 <- glm(Qualityclass_dummy ~ aroma+flavor+acidity, data = data, family = binomial(link = "logit"))
summary(model4)
summ(model4)
####### need a table for summ
```

```{r}
Models<-c('model1','model2','model3','model4')
model.comp.values.model1<-glance(model1)
kable(model.comp.values.model1,digits=2,caption = "model-1 summary")
model.comp.values.model2<-glance(model2)
kable(model.comp.values.model2,digits=2,caption = "model-2 summary")
model.comp.values.model3<-glance(model3)
kable(model.comp.values.model3,digits=2,caption = "model-3 summary")
model.comp.values.model4<-glance(model4)
kable(model.comp.values.model4,digits=2,caption = "model-4 summary")
```

```{r}
model_table <- bind_rows(
  data.frame(Model = "model1", AIC = AIC(model1, k = 2), BIC = BIC(model1)),
  data.frame(Model = "model2", AIC = AIC(model2, k = 2), BIC = BIC(model2)),
  data.frame(Model = "model3", AIC = AIC(model3, k = 2), BIC = BIC(model3)),
  data.frame(Model = "model4", AIC = AIC(model4, k = 2), BIC = BIC(model4))
)

kable(model_table, digits = 2, caption = "Comparison for the 4 models")
```

-   It shows aroma, flavor, acidity, altitude_mean_meters are both significant these varibales will be saved and AIC decreased to min.

# (Analysis) optimization model (final model)

$$\ln\left(\frac{p}{1-p}\right) = \alpha + \beta_1 \cdot \textrm{aroma} +
    \beta_2 \cdot \textrm{flavor}+
    \beta_3 \cdot \textrm{acidity}+ 
    \beta_4 \cdot \textrm{altitudes}$$

-   (elements explains)

```{r}
levels(data$Qualityclass) #base on "good"
#for original model
mod1coefs1 <- round(coef(model1), 2)
library(knitr)
confint(model1) %>%
  kable(caption = "Confidence Interval of model-1")
#for optimization model
mod1coefs3 <- round(coef(model3), 2)
library(knitr)
confint(model3) %>%
  kable(caption = "Confidence Interval of model-3")
```

### log-odds

```{r}
mod.coef.logodds<-model3 %>%
                  summary() %>%
                  coef()
data<- data%>%
      mutate(logodds.good = predict(model3))
```

```{r}
#| label: fig-4
#| fig-cap: the log-odds of explanatory variables for quality good
#| fig-align: left
plot_model(model3, show.values = TRUE, transform = NULL,
           show.p = FALSE)
```

-   Log-odds plot shows.....

### odds

```{r}
model3 %>%
 coef() %>%
  exp()
#check value
exp(coef(model3))
data<- data%>%
      mutate(odds.good = exp(logodds.good))
data<- data%>%
      mutate(prob.good = fitted(model3))
```

```{r}
#| label: fig-5
#| fig-cap: the odds of explanatory variables for quality good
#| fig-align: left
#odd ratio for quality good
plot_model(model3, show.values = TRUE, axis.lim=c(0,10000),
           show.p = FALSE)
```

-   We interpret the odds ratios as follows:....

### probability continuous

```{r}
#aroma/acidity/flavor prob
data_long1 <- data %>%
 pivot_longer(cols = c(aroma, flavor,acidity), names_to = "Type", values_to = "Value")
```

```{r}
#| label: fig-6
#| fig-cap: the prob of aroma/acidity/flavor for quality good
ggplot(data = data_long1, aes(x =Value, y =prob.good, color = Type)) +
 geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
 labs(x = "score", y = "Probability of quality being good", color = "character") +
 theme_minimal()
```

```{r}
#| label: fig-7
#| fig-cap: the prob of altitude_mean_meters for quality good
ggplot(data = data, aes(x =altitude_mean_meters, y =prob.good)) +
 geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
 labs(x = "kilometers", y = "Probability of quality being good", color = "character") +
 theme_minimal()
```

```{r}
#| label: fig-8
#| fig-cap: Probility of quality being good
#| fig-width: 6
#| fig-height: 5
#| fig-align: center
p1=plot_model(model3, type = "pred",terms = "aroma" ,title = "Aroma",
            axis.title = c("aroma", "Prob. of quality being good"))
p2=plot_model(model3, type = "pred", terms="flavor",title = "Flavor",
            axis.title = c("flavor", "Prob. of quality being good"))
p3=plot_model(model3, type = "pred",terms = "acidity", title = "Acidity",
            axis.title = c("acidity", "Prob. of quality being good"))
p4=plot_model(model3, type = "pred",terms = "altitude_mean_meters", title = "altitude_mean_meters",
            axis.title = c("altitude_mean_meters", "Prob. of quality being good"))
#merge
grid.arrange(p1,p2,p3,p4,nrow=2)
```

# Conclusions {#sec-Conc}

-   1 Finally, we choose Model 3....

-   2 Through the log-odds and odds ratio plots, flavor was the most important factor in this research....

-   3....

-   4....
